Cybersecurity Is Too Important\\To Be Left To The Specialists

### 2019-06-09 Response.tex

\font\manual=manfnt
\def\dbend{{\manual\char127}}
\def\eqq{{\buildrel?\over=}}
\def\C{\mathbf{ C}}
\def\N{\mathbf{ N}}
\def\Q{\mathbf{ Q}}
\def\R{\mathbf{ R}}
\def\Z{\mathbf{Z}}
\def\F{\mathbf{F}}
\def\softO{\tilde{\cal O}}
\def\O{{\cal O}}
%\def\N{{\bf N}}
%\def\Q{{\bf Q}}
%\def\R{{\bf R}}
%\def\Z{{\bf Z}}
\def\action#1{\hfil\rlap{\bf#1}\break}
\def\pcite#1{[#1]}
\def\Res{\mathop{\rm Res}\nolimits}
\def\r#1#2{``#1''$\rightarrow$``#2''}
\documentclass{article}
\bibliographystyle{alphaurl}
\usepackage[hyphens]{url}
%\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{verbatim}
\usepackage{hyperref}
\usepackage[show]{ed}
\usepackage{graphicx}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{notation}{Notation}
\newtheorem{example}{Example}
\newtheorem{problem}{Problem}
\def\decision#1{\par{\bf #1}}
\pagestyle{empty}
\begin{document}
\author{Crick/Davenport/Irons/Prickett}
\title{Response to Referees}
\date{June 2019}
\maketitle

\section{Referee 1}
No comments that require addressing in detail.
\section{Referee 2}
No comments that require addressing in detail.
\section{Referee 3}
\begin{enumerate}
\item ``The research appears to be poorly structured and the analysis/argument is hard to interpret''.
\item[A]We have done what we can, but note that the other referees disagree.
\item ``The abstract is too long and could be made concise. For example, the 2nd paragraph can be avoided as the main idea has been conveyed in the 1st paragraph as well.''
\item[A]
\item 
\item[A]
\item 
\item[A]
\item 
\item[A]
\item 
\item[A]
\end{enumerate}
\section{Programme Committee}
No comments that require addressing in detail.

\end{document}

\end{document}
\bibliography{../../../../jhd}
\section{}
\begin{description}
\item
\end{description}
\begin{itemize}
\item
\end{itemize}
\begin{example}
\end{example}
\begin{definition}
\end{definition}
\begin{theorem}
\end{theorem}
\begin{enumerate}
\end{enumerate}
\begin{description}
\item[Theme 1]
\item[Theme 2]
\item[Theme 3]
\item[Theme 4]
\item[Theme 5]
\end{description}

### 2019-06-09 FIE2019-final.tex
JHD starts the process

\begin{IEEEkeywords}
TBC
\end{IEEEkeywords}

@@

Cybersecurity has been in the news for several years, generally prompted by spectacular breaches of one kind or another, such as \cite{BritishAirways2018a}.
It needs attention across the spectrum:
\begin{quote}
change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation. \cite{Hannigan2019a}\footnote{Former Director of GCHQ: British equivalent of NSA.}
\end{quote}

There are calls for education to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}. There has been a recent international working group \cite{Parrishetal2018a}, but this has yet to report.  We note that part of its third aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these (see section \ref{sec:SQL}).

Nevertheless, it is one thing to write curricula and requirements, and
another thing to deliver appropriate education, and one could
reasonably ask how well this is done in practice.
\subsection{CyberSecurity: for all or for specialists?}
In one sense, this title is a false dichotomy: there is a serious need for Cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} Cybersecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.

This is not a brand-new concern: see \cite{Parr2014a} for concerns over five years ago, but it is a growing one. On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of Cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''. It is hard to envisage one person possessing all these attributes, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing boards to invest more in Cybersecurity. On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for staff in this area.
% JHD what are these?  Also the next paragraph needs thinking about - TP - I added the next paragraph but now wonder if it should be deleted - I wonder if there is a point to be made about ''privacy by design and ''privacy by default'' but agile is covered better later.

@@

Specialist curricula abound: a recent one is \cite{ACMIEEEAISSIGSECIFIP}. However, given the shortage of specialists, in practice a good generalist with some Cybersecurity expertise can often get a ``specialist'' job.
\subsection{Research Questions}
There are various levels of specialism at which Cybersecurity education can be addressed.
\begin{enumerate}
\item The person in the street --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
% ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry? Minor point but maybe worth mentioning e.g. being safe online, accessing public services, etc}})
%\item[JHD]This is important, agreed, but I think this is being a user: think car driving rather than car design. BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
\item The general CS graduate.
\item The general CS masters graduate.
\item The specialist CS graduate.
\item The specialist CS masters graduate.
\end{enumerate}
The focus of this paper is on 2--3: the general CS graduate.
\begin{description}
\item[RQ1]What should be taught to the generalist, and how?
\item[RQ2]Should this be taught stand-alone or integrated? 
\item[RQ3]How might accreditation regard Cybersecurity Education, or help with it?
\end{description}
\cite[p.~97]{ACM2013a} takes a distinct view on RQ2:
\begin{quote}
The Information Assurance and Security KA is unique among the set of KAs presented here
given the manner in which the topics are pervasive throughout other Knowledge Areas.
\end{quote}
It proposes 9 ``core'' hours and 63.5 distributed across the other Knowledge Areas.
%\pagebreak - to fix URL issues
\par
Nevertheless, the situation on the ground is different: \cite{Ackerman2019a}, describing the USA, writes as follows
\begin{quote}
Universities suffer shortcomings, as well. Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity. There is a big catch, however. Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.
\end{quote}

@@

It is at least plausible to attribute this difference to differences in the accreditation regimes.
\begin{description}
\item[UK]BCS has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed Cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle). More precisely since 2010, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced Cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
%BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
\item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body. ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots. This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
%\cite[chapter 7]{Davenportminor2019a}.  % need a better reference
\end{description}
\section{Policy Context}

% TC to add policy+skills context
% inc. new EU Cybersecurity Act: https://ec.europa.eu/commission/news/cybersecurity-act-2018-dec-11_en

There has been substantial CS/digital curriculum reform across the
UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,wgictreview:2013,brown-et-al:toce2014,moller+crick:jce2018}
--- but what has happened to the focus on Cybersecurity?  The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention Cybersecurity, but merely as a sub-item in one work package (1.2a).

@@

There is large-scale media
attention on the ``Cybersecurity skills gap''
e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
and
\url{https://www.itpro.co.uk/cyber-security/31554/uk-government-lacks-urgency-in-tackling-cyber-security-skills-gap}
etc -- link to UK Digital Strategy and UK Industrial Strategy.

@@

Is teaching Cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a}).  % JHD - need's fixing.

@@

%Alastair to write about their Credit Fraud experience (MSc Cybersecurity).  HE Academy grant (to be referenced).

@@

Clearly such a major weakness should be well-taught. One would like to think it is, but the first author has yet to meet a UK undergraduate who recalls being taught it \emph{outside} a specialist Cybersecurity course.  In general it's hard to determine what is taught, but a reasonable proxy for this is the content of recomended textbooks.

@@

\subsection{The case of Java}\label{sec:Java}
To the best of the authors' knowledge, no survey equivalent to \cite{Drop2019} has been done for Java textbooks. Indeed, many such books go nowhere near security applications.  But this means that the programmer who has to implement security is left to the documentation of the package/API being used, and to informal resources. \cite{Mengetal2018a} analysed 503 Cybersecurity-related postings on the popular Stack overflow (\url{https://stackoverflow.com}) resource.  53\% were about the Spring Security framework (\url{https://projects.spring.io/spring-security/}), dominated by authentication (45\%). The discussion \cite[\S4.3.1]{Mengetal2018a} of cross-site request forgery (CSRF) is especially worrying.  By default, Spring implcitly enables protection against this. But all the accepted answers to CSRF-related failures simply suggested disabling they check. There were no negative comments about this, and indeed a typical response is ``Adding \verb!csrf().disable()!
solved the issue!!! I have no idea why it was enabled by default''. As of writing (16 January 2018) there were no negative comments about this disabling of a vital security feature.
\par
This research was further developed by \cite{Chenetal2019a}  (and popularised in a security community in \cite{Zorz2019a}). Their first finding was
\begin{quote}
644 out of the 1,429 inspected answer posts
(45\%) are insecure, meaning that insecure suggestions
popularly exist on SO. Insecure answers dominate, in
particular, the SSL/TLS category [355 insecure versus 150 secure, i.e. $>70$\%].
\end{quote}

\subsection{Android}\label{sec:Android}
\cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications. Their key finding was this.
\begin{quote}
We found that 15.4\% of all 1.3 million Android applications
contained security-related code snippets from
Stack Overflow. Out of these 97.9\% contain at least one
insecure code snippet.
\end{quote}
We should note two caveats (in opposite directions). Their labelling was conservative, in that snippets were only labelled as insecure if that was demonstrable, and, for example, mere use of outdated SSL/TLS was not automatically deemed insecure. On the other hand, the insecure snippet might have been used in a way that did not expose the insecurity.
\par
The uncritical reading of Stack Overflow was also noted in \cite[Slide 29]{Votipkaetal2019a}. Their key recommendation  \cite[Slide 32]{Votipkaetal2019a} is ``Improve documentation: Clarify what you can(not) copy/paste''. 
\subsection{Agile}
To the best of the authors' knowledge, no survey equivalent to \cite{Drop2019} has been done for ``Agile'' textbooks. It is the authors' experience that the vast majority of students' individual software projects are developed on Agile lines. Many authors have found disconnects between Agile practices and secure software development: notably \cite{Bartsch2011a} for small projects and \cite{vanderHeijden:2018:EPS:3239235.3267426} for large projects.
\par
Agile's preference for functionality over non-functional requirements is clearly displayed in practice. \cite{Naiakshinaetal2017a} asked 20 student developers to imagine they
were part of a team working on creating a social networking site for
our university and to implement a password storage mechanism for this. 10 (``primed'') were explicitly told that the storage had to be secure and 10 (``unprimed'') were not. None of the unprimed ones implemented any security.
\subsection{Staff}
It is well known that Cybersecurity skills are in short supply, for example \cite{Page2018a}.
\begin{quote}
% JHD: I am in touch with ESG to try to find more, and someUK-specific numbers.
Research into the state of IT conducted annually by ESG\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years. In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills. This had climbed to 51\% at the beginning of this year. Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.
\end{quote}
The ESG survey is international, but ESG have confirmed that the UK figures are very similar.
%JHD possibly to write about his recruitment experiences.  ? more IoC input?

In the authors' experience, it is proving very difficult to recruit academic staff with specialisms in Cybersecurity. The demand for Cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations. As universities expand their Cybersecurity provision it is not uncommon to find multiple jobs advertised at the same time.  A recent example had a professor of Cybersecurity, two senior academic positions and 2 junior academic positions in one advert. There are examples in the UK of Cybersecurity lecturing jobs remaining unfilled for longer than a year. There are also examples of  Cybersecurity subject groups moving en masse from one university to another.

@@

For a few years now there has been a recognized need in the UK to build knowledge, skills and capacity in the area of Cybersecurity. This need has led to the establishment of a number of initiatives from a number of national governments for example The UK Cyber Security Strategy. \cite{UKCabinetOffice} or National Initiative for Cybersecurity Education (NICE) in the USA \cite{NICE}.

The teaching of Cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of Cybersecurity as part of the Computer Science discipline for a number of years \cite{Hentea2006}. There has  been a debate as to whether Cybersecurity is distinct discipline from Computer Science \cite{McGettreick2013}. The consensus now is that Cybersecurity is both a discipline in its own right and that Cybersecurity should be taught within Computer Science and related degrees. There have been a number of international initiatives international to define curricula to support this for example Computer Science Programmes  \cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised Cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.

In the United Kingdom, Higher Education provision addresses both approaches. A significant number of undergraduate and postgraduate programmes are available in both the areas of Computer Science and Cybersecurity (and closely related fields Computer Security, Digital / Computer Forensics, etc). In the UK, Universities and Colleges Admissions Service (UCAS) lists over 40 Higher Education Institutes (HEI) providing Undergraduate qualifications related to Cybersecurity for entry in September 2019. An even larger number of HEI provide study opportunities related to more general Computer Science. UCAS lists 246 provides for undergraduate programmes related to Computer Science.

Accreditation has evolved to directly addresses the Cybersecurity challenges in both general Computer Science programmes and specialist Cybersecurity programmes. In the UK, accreditation in the broad computing area is being performed by a few different agencies. These include:

\textbf{1. Not for Profit Organisations}
Tech Partnership Degrees is a Not For Profit Organisation that provides endorsements to Higher Education programmes with specific curricula elements aimed at job market requirements. One of the curricula elements is related to Cybersecurity. Tech Partnership Degrees have a specialist scope, endorsing programmes in the area of IT Management for Business and Software Engineering for Business.  Tech Partnership Degrees currently endorse 14 IT Management for Business Programmes and 5 Software Engineering for Business Programmes. As such Tech Partnership Degrees currently have limited impact upon more general Computer Science education and none upon specialist Cybersecurity education.

The Institute of Coding is a not for profit organisation that intends to enhance how Digital Skills are developed in Higher Education in the UK \cite{Davenportetal2019a}. This is likely to include Cybersecurity related skills. Like the Tech Partnership the focus is upon job market requirements. Additionally the Institute is looking at potentially endorsing the demonstrable capabilities of graduates as shown by their university studies, work experience and work placements. Given the size of this initiative this potentially has a significant role to play however at time of writing it is clearly a work in progress.

\textbf{2. National Cyber Security Centre (NCSC)}
The NCSC is a UK Government organisation tasked with enhancing the Cybersecurity of the UK. The NCSC publishes and accredits to a number of Cybersecurity standards \cite{NCSC2018a}. These standard are linked to the ACM recommendations for curricula. To date the major focus of NSSC acreditation has been upon Masters degrees specializing in Cybersecurity. More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in Cybersecurity and Computer Science degrees with a significant Cybersecurity focus \cite{NCSC2018b}.

Currently accredited are 15 Cybersecurity MSc programmes with a further 11 provisional accredited, 3 integrated masters Cybersecurity programmes and 1 Cybersecurity Degree Programme with a further 2 provisionally accredited. Hence the extent of accreditation is currently reasonably limited in terms of reach to Computer Science programmes. This appears to be a positive initiative that will hopefully further develop over time.

\textbf{3. Professional Bodies}
The BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of Computer Science and the more specialist area of Cybersecurity discipline areas. The accreditation provided by these institutes are underpinned by international Initiatives such as the Washington Accord\footnote{\url{http://www.ieagreements.org/accords/washington/}.} and Seoul Accord\footnote{\url{https://www.seoulaccord.org/}.}. These memoranda support the internationalising of the curriculum and promote consistency and parity in Computer Science Education globally.   These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006}, \cite{Mutereko2017}. Both the IET and The BCS have a long history of expecting coverage environmental factors within the programmes they accredit. The BCS has for a number of years been expecting significant coverage of Legal, Ethical, Social and Professional Issues \cite{Brooke2018}. Clearly Cybersecurity has been and continues to be part of these expectations.

>>>>

\begin{IEEEkeywords}
Cybersecurity, accreditation, curricula, computer science education, UK
\end{IEEEkeywords}

@@

Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches of one kind or another, such as \cite{BritishAirways2018a}.
It needs attention across the spectrum:
\begin{quote}
change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation. \cite{Hannigan2019a}\footnote{Former Director of GCHQ: British equivalent of NSA.}
\end{quote}

With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}; this is alongside a wider focus on computer science education reform, especially across the UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,brown-et-al:toce2014,wgictreview:2013,moller+crick:jce2018,davenport-et-al:latice2016,murphy-et-al:programming2017,simon-et-al:sigcse2018,tryfonas+crick:petra2018}. There has been a recent international working group \cite{Parrishetal2018a}, but this has yet to report.  We note that part of its third aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these (see section \ref{sec:SQL}).

Nevertheless, it is one thing to write national curricula, specifications and requirements, and
another thing to deliver appropriate and relevant education and skills; furthermore, one could
reasonably ask how well this is done in practice.

\subsection{CyberSecurity: for all or for specialists?}
In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} cybersecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.

This is not a brand-new concern: see \cite{Parr2014a} for concerns over five years ago, but it is a growing one. On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''. It is hard to envisage one person possessing all these attributes, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing boards to invest more in cybersecurity. On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for staff in this area.
% JHD what are these?  Also the next paragraph needs thinking about - TP - I added the next paragraph but now wonder if it should be deleted - I wonder if there is a point to be made about ''privacy by design and ''privacy by default'' but agile is covered better later.

@@

Specialist curricula abound: a recent one is \cite{ACMIEEEAISSIGSECIFIP}. However, given the shortage of specialists, in practice a good generalist with some cybersecurity expertise can often get a ``specialist'' job.
\subsection{Research Questions}
There are various levels of specialism at which cybersecurity education can be addressed.
\begin{enumerate}
\item The person in the street --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
% ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry? Minor point but maybe worth mentioning e.g. being safe online, accessing public services, etc}})
%\item[JHD]This is important, agreed, but I think this is being a user: think car driving rather than car design. BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
\item The general CS graduate.
\item The general CS masters graduate.
\item The specialist CS graduate.
\item The specialist CS masters graduate.
\end{enumerate}
The focus of this paper is on 2--3: the general CS graduate.
\begin{description}
\item[RQ1]What should be taught to the generalist, and how?
\item[RQ2]Should this be taught stand-alone or integrated? 
\item[RQ3]How might accreditation regard cybersecurity Education, or help with it?
\end{description}
\cite[p.~97]{ACM2013a} takes a distinct view on RQ2:
\begin{quote}
The Information Assurance and Security KA is unique among the set of KAs presented here
given the manner in which the topics are pervasive throughout other Knowledge Areas.
\end{quote}
It proposes 9 ``core'' hours and 63.5 distributed across the other Knowledge Areas.
%\pagebreak - to fix URL issues
\par
Nevertheless, the situation on the ground is different: \cite{Ackerman2019a}, describing the USA, writes as follows
\begin{quote}
Universities suffer shortcomings, as well. Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity. There is a big catch, however. Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.
\end{quote}

@@

It is at least plausible to attribute this difference to differences in the accreditation regimes.
\begin{description}
\item[UK]BCS has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle). More precisely since 2010, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
%BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
\item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body. ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots. This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
%\cite[chapter 7]{Davenportminor2019a}.  % need a better reference
\end{description}
\section{Policy Context}

% TC to add policy+skills context
% inc. new EU cybersecurity Act: https://ec.europa.eu/commission/news/cybersecurity-act-2018-dec-11_en

There has been substantial CS/digital curriculum reform across the
UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,wgictreview:2013,brown-et-al:toce2014,moller+crick:jce2018}
--- but what has happened to the focus on cybersecurity?  The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention cybersecurity, but merely as a sub-item in one work package (1.2a).

@@

There is large-scale media
attention on the ``cybersecurity skills gap''
e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
and
\url{https://www.itpro.co.uk/cyber-security/31554/uk-government-lacks-urgency-in-tackling-cyber-security-skills-gap}
etc -- link to UK Digital Strategy and UK Industrial Strategy.

@@

Is teaching cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a}).  % JHD - need's fixing.

@@

%Alastair to write about their Credit Fraud experience (MSc cybersecurity).  HE Academy grant (to be referenced).

@@

Clearly such a major weakness should be well-taught. One would like to think it is, but the first author has yet to meet a UK undergraduate who recalls being taught it \emph{outside} a specialist cybersecurity course.  In general it's hard to determine what is taught, but a reasonable proxy for this is the content of recomended textbooks.

@@

\subsection{The case of Java}\label{sec:Java}
To the best of the authors' knowledge, no survey equivalent to \cite{Drop2019} has been done for Java textbooks. Indeed, many such books go nowhere near security applications.  But this means that the programmer who has to implement security is left to the documentation of the package/API being used, and to informal resources. \cite{Mengetal2018a} analysed 503 cybersecurity-related postings on the popular Stack overflow (\url{https://stackoverflow.com}) resource.  53\% were about the Spring Security framework (\url{https://projects.spring.io/spring-security/}), dominated by authentication (45\%). The discussion \cite[\S4.3.1]{Mengetal2018a} of cross-site request forgery (CSRF) is especially worrying.  By default, Spring implicitly enables protection against this. But all the accepted answers to CSRF-related failures simply suggested disabling they check. There were no negative comments about this, and indeed a typical response is ``Adding \verb!csrf().disable()!
solved the issue!!! I have no idea why it was enabled by default''. As of writing (16 January 2018) there were no negative comments about this disabling of a vital security feature.
\par
This research was further developed by \cite{Chenetal2019a}  (and popularised in a security community in \cite{Zorz2019a}). Their first finding was
\begin{quote}
644 out of the 1,429 inspected answer posts
(45\%) are insecure, meaning that insecure suggestions
popularly exist on SO. Insecure answers dominate, in
particular, the SSL/TLS category [355 insecure versus 150 secure, i.e. $>70$\%].
\end{quote}

\subsection{Android}\label{sec:Android}
\cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications. Their key finding was this.
\begin{quote}
We found that 15.4\% of all 1.3 million Android applications
contained security-related code snippets from
Stack Overflow. Out of these 97.9\% contain at least one
insecure code snippet.
\end{quote}
We should note two caveats (in opposite directions). Their labelling was conservative, in that snippets were only labelled as insecure if that was demonstrable, and, for example, mere use of outdated SSL/TLS was not automatically deemed insecure. On the other hand, the insecure snippet might have been used in a way that did not expose the insecurity.
\par
The uncritical reading of Stack Overflow was also noted in \cite[Slide 29]{Votipkaetal2019a}. Their key recommendation  \cite[Slide 32]{Votipkaetal2019a} is ``Improve documentation: Clarify what you can(not) copy/paste''. 
\subsection{Agile}
To the best of the authors' knowledge, no survey equivalent to \cite{Drop2019} has been done for ``Agile'' textbooks. It is the authors' experience that the vast majority of students' individual software projects are developed on Agile lines. Many authors have found disconnects between Agile practices and secure software development: notably \cite{Bartsch2011a} for small projects and \cite{vanderHeijden:2018:EPS:3239235.3267426} for large projects.
\par
Agile's preference for functionality over non-functional requirements is clearly displayed in practice. \cite{Naiakshinaetal2017a} asked 20 student developers to imagine they
were part of a team working on creating a social networking site for
our university and to implement a password storage mechanism for this. 10 (``primed'') were explicitly told that the storage had to be secure and 10 (``unprimed'') were not. None of the unprimed ones implemented any security.
\subsection{Staff}
It is well known that cybersecurity skills are in short supply, for example \cite{Page2018a}.
\begin{quote}
% JHD: I am in touch with ESG to try to find more, and someUK-specific numbers.
Research into the state of IT conducted annually by ESG\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years. In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills. This had climbed to 51\% at the beginning of this year. Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.
\end{quote}
The ESG survey is international, but ESG have confirmed that the UK figures are very similar.
%JHD possibly to write about his recruitment experiences.  ? more IoC input?

In the authors' experience, it is proving very difficult to recruit academic staff with specialisms in cybersecurity. The demand for cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations. As universities expand their cybersecurity provision it is not uncommon to find multiple jobs advertised at the same time.  A recent example had a professor of cybersecurity, two senior academic positions and 2 junior academic positions in one advert. There are examples in the UK of cybersecurity lecturing jobs remaining unfilled for longer than a year. There are also examples of  cybersecurity subject groups moving en masse from one university to another.

@@

For a few years now there has been a recognized need in the UK to build knowledge, skills and capacity in the area of cybersecurity. This need has led to the establishment of a number of initiatives from a number of national governments for example The UK Cyber Security Strategy. \cite{UKCabinetOffice} or National Initiative for cybersecurity Education (NICE) in the USA \cite{NICE}.

The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the Computer Science discipline for a number of years \cite{Hentea2006}. There has  been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettreick2013}. The consensus now is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within Computer Science and related degrees. There have been a number of international initiatives international to define curricula to support this for example Computer Science Programmes  \cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.

In the United Kingdom, Higher Education provision addresses both approaches. A significant number of undergraduate and postgraduate programmes are available in both the areas of Computer Science and cybersecurity (and closely related fields Computer Security, Digital / Computer Forensics, etc). In the UK, Universities and Colleges Admissions Service (UCAS) lists over 40 Higher Education Institutes (HEI) providing Undergraduate qualifications related to cybersecurity for entry in September 2019. An even larger number of HEI provide study opportunities related to more general Computer Science. UCAS lists 246 provides for undergraduate programmes related to Computer Science.

Accreditation has evolved to directly addresses the cybersecurity challenges in both general Computer Science programmes and specialist cybersecurity programmes. In the UK, accreditation in the broad computing area is being performed by a few different agencies. These include:

\textbf{1. Not for Profit Organisations}
Tech Partnership Degrees is a Not For Profit Organisation that provides endorsements to Higher Education programmes with specific curricula elements aimed at job market requirements. One of the curricula elements is related to cybersecurity. Tech Partnership Degrees have a specialist scope, endorsing programmes in the area of IT Management for Business and Software Engineering for Business.  Tech Partnership Degrees currently endorse 14 IT Management for Business Programmes and 5 Software Engineering for Business Programmes. As such Tech Partnership Degrees currently have limited impact upon more general Computer Science education and none upon specialist cybersecurity education.

The Institute of Coding is a not for profit organisation that intends to enhance how Digital Skills are developed in Higher Education in the UK \cite{Davenportetal2019a}. This is likely to include cybersecurity related skills. Like the Tech Partnership the focus is upon job market requirements. Additionally the Institute is looking at potentially endorsing the demonstrable capabilities of graduates as shown by their university studies, work experience and work placements. Given the size of this initiative this potentially has a significant role to play however at time of writing it is clearly a work in progress.

\textbf{2. National Cyber Security Centre (NCSC)}
The NCSC is a UK Government organisation tasked with enhancing the cybersecurity of the UK. The NCSC publishes and accredits to a number of cybersecurity standards \cite{NCSC2018a}. These standard are linked to the ACM recommendations for curricula. To date the major focus of NSSC accreditation has been upon Masters degrees specializing in cybersecurity. More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in cybersecurity and Computer Science degrees with a significant cybersecurity focus \cite{NCSC2018b}.

Currently accredited are 15 cybersecurity MSc programmes with a further 11 provisional accredited, 3 integrated masters cybersecurity programmes and 1 cybersecurity Degree Programme with a further 2 provisionally accredited. Hence the extent of accreditation is currently reasonably limited in terms of reach to Computer Science programmes. This appears to be a positive initiative that will hopefully further develop over time.

\textbf{3. Professional Bodies}
The BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of Computer Science and the more specialist area of cybersecurity discipline areas. The accreditation provided by these institutes are underpinned by international Initiatives such as the Washington Accord\footnote{\url{http://www.ieagreements.org/accords/washington/}.} and Seoul Accord\footnote{\url{https://www.seoulaccord.org/}.}. These memoranda support the internationalising of the curriculum and promote consistency and parity in Computer Science Education globally.   These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006}, \cite{Mutereko2017}. Both the IET and The BCS have a long history of expecting coverage environmental factors within the programmes they accredit. The BCS has for a number of years been expecting significant coverage of Legal, Ethical, Social and Professional Issues \cite{Brooke2018}. Clearly cybersecurity has been and continues to be part of these expectations.
